---
title: "Titanic - Decision Trees"
output: html_notebook
---

```{r warning=FALSE, message=FALSE}
library(rpart)
library(rattle)
library(dplyr)
library(RCurl)
library(ggplot2)
```

### 1. Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### 2. Displaying data

```{r}
str(train)
str(test)
```

### 3. Removing NAs

```{r}
train <- train %>% na.omit()
test <- test %>% na.omit()
```

### 4. Converting categorical variables to factors

```{r}
train$Survived <- factor(train$Survived)
train$Pclass <- factor(train$Pclass)
test$Pclass <- factor(test$Pclass)
```

### 5. Visualizing the training data

```{r}
ggplot(train, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Training Data")
```

### 6.Selecting features

```{r}
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 7. Creating the Classification Tree

```{r, fig.height=8}
set.seed(9)
tree <- rpart(formula, data=train, method="class")
fancyRpartPlot(tree, uniform=TRUE, main="Classification Tree")
```

### 8. Cross-Validation 

To examine whether the tree model is over fitting, find the size of tree with the minimum error.

```{r}
plotcp(tree)
tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
printcp(tree)
```

### 9. Pruning the Tree

Prune the over fitting notes.

```{r, fig.height=7}
trim <- tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
ptree<- prune(tree, cp=trim)
fancyRpartPlot(ptree, uniform=TRUE, main="Pruned Classification Tree")
```

### 10. Predicting with the test data

```{r}
predict <- predict(ptree, test, type = "prob")
```

### 11. Visualizing the result

```{r}
test$Survived <- predict[,2]
ggplot(test, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Predictiom with the Test Data ")
```


