---
title: "Bike Sharing: Conditional Inference Trees"
output: 
  html_document: 
    keep_md: yes
---

### Data Fields
**datetime** - hourly date + timestamp

**season** -  1 = spring, 2 = summer, 3 = fall, 4 = winter 

**holiday** - whether the day is considered a holiday

**workingday** - whether the day is neither a weekend nor holiday

**weather** - 1: Clear, Few clouds, Partly cloudy, Partly cloudy

2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist 

3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds 

4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog 

**temp** - temperature in Celsius

**atemp** - "feels like" temperature in Celsius

**humidity** - relative humidity

**windspeed** - wind speed

**casual** - number of non-registered user rentals initiated

**registered** - number of registered user rentals initiated

**count** - number of total rentals

```{r warning=FALSE, message=FALSE}
library(rpart)
library(rpart.plot)
library(rattle)
library(data.table)
library(partykit)
library(RCurl)
```

### 1. Reading files

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Bike%20Sharing/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Bike%20Sharing/test.csv')
test <- read.csv(text = url) 
```

### 2. Feature selection

Since temp and atemp are highly correlated, we can remove one of them from the training data. We will remove atemp in following feature set.

```{r}
train<-train[, !(colnames(train) %in% c('atemp'))]
time <- factor(substring(train$datetime, 12, 20))
train$hour<- as.numeric(substr(time, 1, 2))
train$month <- factor(substring(as.Date(train$datetime), 6, 7))
train$workingday <- factor(train$workingday)

test<-test[, !(colnames(test) %in% c('atemp'))] 
time <- factor(substring(test$datetime, 12, 20))
test$hour<- as.numeric(substr(time, 1, 2))
test$month <- factor(substring(as.Date(test$datetime), 6, 7))
test$workingday <- factor(test$workingday)
```

```{r}
formula <- count ~ month + weather + temp + humidity + hour + workingday + holiday
```

### 3. Conditional Inference Trees

```{r}
set.seed(200)
fit <- ctree(formula, data=train)

# Print the tree
print(fit)
#plot(fit, type = "simple")
```

### 4. Prediction

```{r}
predict <- predict(fit, test)
#plot(predict, type = "simple")
#text(predict)
result <- data.frame(datetime = test$datetime, temp = test$temp, month = test$month, workingday = test$workingday, count=predict)
write.csv(result, file="ctree_result.csv", row.names=FALSE)
```

