---
title: "Object Detection with image.darknet"
output: 
  html_document: 
    keep_md: yes
---

## image.darknet

https://github.com/bnosac/image/blob/master/image.darknet/R/darknet_models.R

## Installation of image.darknet

<code>
install.packages("devtools")
</code>

<code>
devtools::install_github("bnosac/image", subdir = "image.darknet", build_vignettes = TRUE)
</code>

```{r, message=FALSE, warning=FALSE}
library(image.darknet)
```

## Creating a modle

```{r}
yolo_tiny_voc <- image_darknet_model(type = 'detect', 
    model = 'tiny-yolo-voc.cfg', 
    weights = system.file(package='image.darknet', 'models', 'tiny-yolo-voc.weights'), 
    labels = system.file(package='image.darknet', 'include', 'darknet', 'data', 'voc.names'))
```

## The function for object detection

```{r}
detect <- function(input, output) {
  a <- image_darknet_detect(file = input, object = yolo_tiny_voc, threshold = 0.19)
  
  file.copy('predictions.png', output)
  knitr::include_graphics(c(input, output))
}
```

```{r}
dir.create(file.path('.', 'output'), showWarnings = FALSE)
```

## Works with a upsidedown cat.

```{r}
detect('input\\cat-yoga.jpg', 'output\\predictions-cat-yoga.png')
```

## Works well with a little bit complicated scene.

```{r}
detect('input\\riding.jpg', 'output\\predictions-riding.png')
```

## Does not work with a person in white.

```{r}
detect('input\\wedding.jpg', 'output\\predictions-wedding.png')
```

## Does not work with a dark person and a partially covered car.

```{r}
detect('input\\zx.jpg', 'output\\predictions-zx.png')
```

## Can AI understand innovation?

```{r}
detect('input\\cargo-bike-1.jpg', 'output\\predictions-cargo-bike-1.png')
```

## No, Lenin is not a cat!

```{r}
detect('input\\berlin.jpg', 'output\\predictions-berlin.png')
```