---
title: "EEG Brain Wave for Confusion"
output:
  html_notebook: default
  html_document: default
---

### Source

https://www.kaggle.com/wanghaohan/eeg-brain-wave-for-confusion

```{r warning=FALSE, message=FALSE}
library(RCurl)
library(formattable)
library(corrplot)
library(ggfortify)
```

###Read Data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/EEG%20Brain%20Wave%20for%20Confusion/EEG-data.csv')
eeg <- read.csv(text = url) 

url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/EEG%20Brain%20Wave%20for%20Confusion/demographic-info.csv')
demography<-read.csv(text = url) 
```

###Data Exploration

####1. EEG Data

```{r}
str(eeg)
unique(eeg$subject.ID)
unique(eeg$predefined.label)
unique(eeg$Self.defined.label)

#Lower triangular of the correlogram
correlation<-cor(eeg)
corrplot(correlation, type="lower")
```

Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients.

Add significance level to the correlogram

```{r}
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
corrplot(correlation, type="lower", p.mat = cor.mtest(eeg), sig.level = 0.01)
```

In the above figure, correlations with p-value > 0.01 are considered as insignificant. In this case the correlation coefficient values are leaved blank or crosses are added.

####2. Demographic Info

```{r}
formattable(demography, list(gender=formatter("span", style = x ~ ifelse(x == "M", style(color = "blue "), style(color = "red")))), align ="l")
```

###PCA

```{r}
pca <- prcomp(eeg, center = TRUE, scale. = TRUE) 
print(pca)
plot(pca, type = "l")
summary(pca)

eeg$subject.ID<-as.factor(eeg$subject.ID)
eeg$predefined.label<-as.factor(eeg$predefined.label)
eeg$Self.defined.label<-as.factor(eeg$Self.defined.label)

autoplot(pca, data =eeg, colour ='Self.defined.label', alpha=I(0.4))
autoplot(pca, data =eeg, colour ='predefined.label', alpha=I(0.4), loadings = TRUE, loadings.label = TRUE, loadings.colour = 'blue4', loadings.label.colour = 'darkgreen')
autoplot(pca, data =eeg, colour ='subject.ID', alpha=I(0.3))
```
