---
title: "Titanic - Recursive partitioning"
output: html_document
---

### Variable Descriptions

**survival**:  Survival (0 = No; 1 = Yes) 

**pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**name**:      Name 

**sex**:       Sex 

**age**:       Age 

**sibsp**:     Number of Siblings/Spouses Aboard 

**parch**:     Number of Parents/Children Aboard 

**ticket**:    Ticket Number 

**fare**:      Passenger Fare 

**cabin**:     Cabin 

**embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 

```{r warning=FALSE, message=FALSE}
library(gtools)
library(ggplot2)
library(RCurl)
library(rpart)
library(rpart.plot)
library(partykit)
library(dplyr)
```

### 1. Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### 2. Removing names and NAs.

```{r}
train<-train[, !(colnames(train) %in% c('PassengerId', 'Name', 'Ticket', 'Cabin'))]
train <-train %>% na.omit()
dim(train)
test<-test[, !(colnames(test) %in% c('PassengerId', 'Name', 'Ticket', 'Cabin'))]
test <- test %>% na.omit()
dim(test)
```

### 3. Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 4. Classification Tree
Creating tree

```{r}
set.seed(2020)
fit <- rpart(formula, data=train, method="class")
as.party(fit)
rpart.plot(fit, main="default rpart.plot\n(type = 0, extra = 0)")
```

### 5. Validation

```{r}
# Visualize cross-validation results
rsq.rpart(fit)
```

### 5. Prediction

```{r}
predict <- predict(fit, test, type = "prob")
predict
```
