---
title: "Titanic - Data Exploration"
output: 
  html_document: 
    keep_md: yes
---

### Variable Descriptions

**survival**:  Survival (0 = No; 1 = Yes) 

**pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**name**:      Name 

**sex**:       Sex 

**age**:       Age 

**sibsp**:     Number of Siblings/Spouses Aboard 

**parch**:     Number of Parents/Children Aboard 

**ticket**:    Ticket Number 

**fare**:      Passenger Fare 

**cabin**:     Cabin 

**embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 

```{r warning=FALSE, message=FALSE}
library(gtools)
library(ggplot2)
library(caret)
library(dplyr)
library(RCurl)
```

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
dim(train)
```

Removing names and NAs.

```{r}
train<-train[, !(colnames(train) %in% c('name'))]
train <-train %>% na.omit()
```

```{r}
par(mar = rep(2, 4))
trainCols <- c("Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked")
featurePlot(x=train[,trainCols], y=train$Survived, plot='pairs', par.settings=list(superpose.symbol=list(alpha = rep(1, 9), cex=rep(0, 1), font = rep(1, 2), pch=c(0))))

dev.copy(png, width=900, height=900, file="Titanic-features.png")  
dev.off()
```

```{r}
train$Survived <- factor(train$Survived)
train$Sex <- factor(train$Sex)
train$Pclass <- factor(train$Pclass)
```

```{r}
ggplot(train, aes(Fare, Pclass)) +     
  geom_point(stat="identity", alpha=0.1) +
  scale_x_log10() +
  ggtitle("Correlation of Fare and Pclass")
```

```{r}
ggplot(data=train, aes(x=Sex, y=Age, fill=Survived)) +
    geom_violin() +       
    xlab("Sex") +
    ylab("Age") +
    guides(fill=guide_legend(title="Sex")) + 
    ggtitle("Correlation of Age and Sex")
```

```{r}
ggplot(data=train, aes(x=Sex, y=Age, fill=Survived)) +
    geom_violin() +   
    facet_grid(. ~ Survived) +   
    xlab("Sex") +
    ylab("Age") +
    guides(fill=guide_legend(title="Survived")) + 
    ggtitle("Correlation of Survived, Age, and Sex")
```

```{r}
ggplot(data=train, aes(x=Sex, y=Age, fill=Survived)) +
    geom_violin() +   
    facet_grid(Pclass ~ Survived) +   
    xlab("Sex") +
    ylab("Age") +
    guides(fill=guide_legend(title="Survived")) + 
    ggtitle("Correlation of Survived, Age, Sex, and Class")
```

```{r}
ggplot(train, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Training Data")
```

Suggested training set formula: <code>Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked</code>
