---
title: "Titanic: Random Forests"
output: 
  html_document: 
    keep_md: yes
---

### Variable Descriptions

**Survival**:  Survival (0 = No; 1 = Yes) 

**Pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**Name**:      Name 

**Sex**:       Sex 

**Age**:       Age 

**SibSp**:     Number of Siblings/Spouses Aboard 

**Parch**:     Number of Parents/Children Aboard 

**Ticket**:    Ticket Number 

**Fare**:      Passenger Fare 

**Cabin**:     Cabin 

**Embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 
 
```{r warning=FALSE, message=FALSE}
library(caret)
library(dplyr)
library(ggplot2)
library(RCurl)
```

### 1. Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### 2.  Removing NA.

```{r}
train<-train[, !(colnames(train) %in% c('Name', 'Ticket', 'Cabin'))]
train <-train %>% na.omit()
test<-test[, !(colnames(test) %in% c('Name', 'Ticket', 'Cabin'))]
test <- test %>% na.omit()
```

### 3. Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 4. Creating random forests model

```{r}
set.seed(1212)
fit <- suppressMessages(train(formula, data=train, method="rf"))
fit
ggplot(varImp(fit))
ggplot(fit, metric = "Accuracy")
ggplot(fit, metric = "Kappa")
```

### 5. Predicting with the test data set

```{r}
Survived <- predict(fit, test)
r<-cbind(Survived, test)
ggplot(r, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Test Data Predicted")

ggplot(train, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Training Data")

total<-rbind(train, r)
ggplot(total, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Total")
```

