---
title: "Exploring SQL Server Database with R"
output: html_document
---

This project is created for Boston Code Camp 23. It demonstrates how to connect and query the SQL Server database with <code>RODBC</code>, and how to explore and visualize the data with <code>ggplot2</code>. The focus is on the technology rather than the interpretation of the content.  I use AdventureWorks as the sample database because it is free and widely available. Please be aware that the schema is slightly different in each version. You may need to modify the query statement if you find this issue with your installed AdventureWorks database.

For data exploring with <code>ggplot2</code>, I build plots from a simple one, add features incrementally, and achieve relatively sophisticated plots in the end. I also create several identical plots with both <code>qplot</code> and <code>ggplot</code>, respectively. This will be helpful for beginners to better understand the concepts and techniques of <code>ggplot2</code>.
 

### System Requirements
- Microsoft SQL Server 2005 or higher

- AdventureWorks Databases http://msftdbprodsamples.codeplex.com/releases/view/125550. For older versions, http://msftdbprodsamples.codeplex.com/releases/view/93587

- Microsoft ODBC Driver for SQL Server http://www.microsoft.com/en-US/download/details.aspx?id=36434

- R http://cran.r-project.org/

### Demo: Conneting to SQL Server database
Connecting to AdventureWorks database with <code>RODBC</code>.

```{r}
library(RODBC)
conn <- odbcDriverConnect("Driver=ODBC Driver 11 for SQL Server;Server=HP2;Database=AdventureWorks;Trusted_Connection=yes")
```

### Demo: Querying with the stored procedure
The returned result is in a data frame.

```{r}
# AdventureWorks query
#managers<-sqlQuery(conn,"exec dbo.uspGetEmployeeManagers @EmployeeID = 1")
# AdventureWorks2014 query
managers<-sqlQuery(conn,"exec dbo.uspGetEmployeeManagers @BusinessEntityID = 5")
head(managers)
```
###Demo: Generating a simple table with knitr
Displaying the data with <code>kable</code>, a simple table generator.

```{r}
library(knitr)
kable(managers)
```

### Demo: Querying with the SQL statement
Querying the sales data. The returned result is in a data frame.

```{r}
sales <- sqlQuery(conn, "SELECT soh.OrderDate,                                            
                        cr.Name as CountryRegion,
                        p.Name AS ProductName,                                           
                        soh.TotalDue,
                        ps.Name AS Category,
                        pc.Name AS Subcategory,
                        CASE WHEN soh.OnlineOrderFlag >0 THEN 'TRUE' ELSE 'FALSE' END AS OnlineOrder
                        FROM [AdventureWorks].[Sales].[SalesOrderHeader] soh
                        INNER JOIN [AdventureWorks].[Sales].[SalesTerritory] AS t 
                            ON t.territoryid=soh.territoryid
                        INNER JOIN [AdventureWorks].[Person].[CountryRegion] AS cr 
                            ON cr.countryregioncode=t.countryregioncode
                        INNER JOIN [AdventureWorks].[Sales].[SalesOrderDetail] AS sod
                            ON soh.SalesOrderID = sod.SalesOrderID
                        INNER JOIN [AdventureWorks].[Production].[Product] AS p
                            ON sod.ProductID = p.ProductID
                        INNER JOIN [AdventureWorks].[Production].[ProductSubcategory] AS ps
                            ON p.ProductSubcategoryID = ps.ProductSubcategoryID
                        INNER JOIN [AdventureWorks].[Production].[ProductCategory] AS pc
                            ON ps.ProductCategoryID = pc.ProductCategoryID
                        ORDER BY cr.Name")
close(conn)
head(sales, 3)
str(sales)
summary(sales)
```

###Demo: Exploring the sales history
This section shows how to generate the same plots with both <code>qplot</code> and <code>ggplot</code> step by step. Let's start with qplot and add features incrementally.

```{r}
library(ggplot2)
qplot(data=sales, y=TotalDue, x=OrderDate, main='Simple Plot (qplot)')
qplot(data=sales, y=TotalDue, x=OrderDate, color=OnlineOrder, size=TotalDue, main='Plot with aesthetic mappings (qplot)')
qplot(data=sales, y=TotalDue, x=OrderDate, color=OnlineOrder, size=TotalDue, main='Plot with log and aesthetic mappings (qplot)') + scale_y_log10()

```

Now let's generate the same plot set with <code>ggplot</code> step by step.

```{r}
ggplot(sales, aes(OrderDate, TotalDue)) + geom_point() + ggtitle('Simple Plot (ggplot)')
ggplot(sales, aes(OrderDate, TotalDue)) + geom_point(aes(colour = OnlineOrder, size=TotalDue)) + ggtitle('Plot with aesthetic mappings (ggplot)')
ggplot(sales, aes(OrderDate, TotalDue)) + geom_point(aes(colour = OnlineOrder, size=TotalDue)) + ggtitle('Plot with log and aesthetic mappings (ggplot)') + scale_y_log10()
```

###Demo: Bar plot with qplot and ggplot
This section shows how to create bar plots with <code>qplot</code> and <code>ggplot</code>. Please also pay attention to how to use <code>lubridate</code> to extract the month value from OrderDate.

```{r}
library(lubridate)
qplot(month(OrderDate), TotalDue/1000000, fill=CountryRegion, data=sales, geom="bar", stat="identity", xlab="Month", ylab="Value ($m)",  main='Bar Plot (qplot)') +
   scale_x_discrete(limits=c(1:12)) 
```

```{r}
ggplot(data=sales,aes(x=month(OrderDate),  y=TotalDue/1000000, fill=CountryRegion)) + 
  geom_bar(stat="identity") + 
  labs(x="Month", y="Value ($m)") +
  scale_fill_brewer(type="qual") +
  scale_x_discrete("Month", limits=c(1:12)) +
  ggtitle("Bar Plot (ggplot)")
```

The above two bar plots should be the same, except the built-in pallet.

###Demo: Bar plot with flipped cartesian coordinates
Following two examples show how to use <code>coord_flip</code> to flip the cartesian coordinates. Please also pay attention to how to convert OrderDate to year with <code>lubridate</code>.

```{r}
qplot(year(OrderDate), TotalDue/1000000, fill=CountryRegion, data=sales, geom="bar", stat="identity", xlab="Year", ylab="Value ($m)", main="Bar Plot with Coordinate Flip (qplot)") + coord_flip()
```


```{r}
ggplot(data=sales,aes(x=year(OrderDate), y=TotalDue/1000000, fill=CountryRegion)) + 
  geom_bar(stat="identity") +  
  labs(x="Year",y="Value ($m)") +
  scale_fill_brewer(type="qual") +
  ggtitle("Bar Plot with Coordinate Flip (ggplot)") +
  coord_flip()
```

The above two bar plots should be identical, except the built-in pallet.

###Demo: Tabular display
The tabular plot is excellent in representing multivariate data. Let's start with displaying sales per year per country. 

```{r}
ggplot(sales, aes(OrderDate, TotalDue)) + 
  geom_point(aes(color=factor(CountryRegion), fill = factor(CountryRegion))) + 
  facet_grid(.~CountryRegion) + 
  theme(legend.position="none", axis.text.x=element_text(angle=90))
```

It appears there are a lot of overlapped data points; and we can't see how many data points are overlapped. We definitely need to improve the plot. Let's try to use the violin plot to show the density of data.

```{r}
ggplot(sales, aes(OrderDate, TotalDue)) + 
  geom_violin(aes(fill = factor(CountryRegion))) + 
  facet_grid(.~CountryRegion) + 
  theme(legend.position="none", axis.text.x=element_text(angle=90))
```

Looks good! Let's overlap the points with violin in the same plot. The order matters!

```{r}
ggplot(sales, aes(OrderDate, TotalDue)) + 
  facet_grid(.~CountryRegion) + 
  theme(legend.position="none", axis.text.x=element_text(angle=90)) + 
  geom_point(aes(color=factor(CountryRegion), fill = factor(CountryRegion))) +
  geom_violin(aes(fill = factor(CountryRegion)))

```

Looks pretty cool! We have learned how to overlap two plots together for richer representation. 

Depending on characteristics of data, sometimes you may need to try different plots to better represent the density. As an exercise, you can replace the <code>geom_violin</code> with either <code>geom_boxplot</code> or <code>geom_density</code> to see the different plots. You don't know which one is the best for your data until you try.

###Demo: Lay out panels in a grid
Let's generate an arrry of bar plots in the panel. Each plot is associated to a country. Note how to set the palette with the index.

```{r}
ggplot(data=sales,aes(x=year(OrderDate),y=TotalDue/1000000, fill=CountryRegion)) +
  geom_bar(stat="identity") + 
  facet_grid(.~CountryRegion) +
  theme(legend.position="none",axis.text.x=element_text(angle=90)) +
  labs(x="Year",y="Value ($m)") +
  scale_fill_brewer(type="qual", palette=3)
```

Finally, let's format the above plot with the Subcategory. Note how to set the palette with the name.

```{r}
#library(lubridate)
ggplot(data=sales,aes(x=year(OrderDate), y=TotalDue/1000000, fill=CountryRegion)) +
  geom_bar(stat="identity") +
  facet_grid(Subcategory~CountryRegion) +
  theme(legend.position="none", axis.text.x=element_text(angle=90)) + 
  labs(x="Year",y="Value ($m)") +
  scale_fill_brewer(type="qual", palette='Paired')
```
