---
title: "Titanic - Decision Trees"
output: 
  html_document: 
    keep_md: yes
---

### Variable Descriptions

**survival**:  Survival (0 = No; 1 = Yes) 

**pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**name**:      Name 

**sex**:       Sex 

**age**:       Age 

**sibsp**:     Number of Siblings/Spouses Aboard 

**parch**:     Number of Parents/Children Aboard 

**ticket**:    Ticket Number 

**fare**:      Passenger Fare 

**cabin**:     Cabin 

**embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton)  

```{r warning=FALSE, message=FALSE}
library(rpart)
library(rattle)
library(rpart.plot)
library(dplyr)
library(RCurl)
```

### Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### Removing NA.

```{r}
train<-train[, !(colnames(train) %in% c('name'))]
train <-train %>% na.omit()
test<-test[, !(colnames(test) %in% c('name'))]
test <- test %>% na.omit()
```

### 1. Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 2. Classification Tree
Creating tree

```{r}
fit <- rpart(formula, data=train, method="class")
fancyRpartPlot(fit)
```

### prune tree
Prune back the tree to avoid overfitting the data. Typically, you will want to select a tree size that minimizes the cross-validated error, the xerror column printed by printcp( ).

```{r}
# prune the tree 
pfit<- prune(fit, cp=fit$cptable[which.min(fit$cptable[,"xerror"]),"CP"])
fancyRpartPlot(pfit)
```

### Prediction

```{r}
Prediction <- predict(pfit, test, type = "prob")
#Prediction
```

### Trying to tweak rpart.control control parameters.

```{r}
fit <- rpart(formula, data=train, method="class", control=rpart.control(minsplit=2, cp=0))
fancyRpartPlot(fit)

fit <- rpart(formula, data=train, method="class", control=rpart.control(minsplit=2, cp=0, minbucket=2))
fancyRpartPlot(fit)
```

### 3. Regression Tree

```{r}
fit <- rpart(formula, data=train, method="anova")
# Plot tree 
fancyRpartPlot(fit)
# Visualize cross-validation results
plotcp(fit) 
# Detailed summary of splits
summary(fit)
```

### 4. Random Forests

Constructing a multitude of decision trees at training time. Random decision forests correct for decision trees' habit of overfitting to their training set.

```{r}
library(randomForest)
fit <- randomForest(formula, data=train)
print(fit)
importance(fit)
```


