---
title: "Titanic: Machine Learning from Disaster - GBM (Gradient Boosting Machine)"
output: 
  html_document: 
    keep_md: yes
---


### Variable Descriptions

**survival**:  Survival (0 = No; 1 = Yes) 

**pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**name**:      Name 

**sex**:       Sex 

**age**:       Age 

**sibsp**:     Number of Siblings/Spouses Aboard 

**parch**:     Number of Parents/Children Aboard 

**ticket**:    Ticket Number 

**fare**:      Passenger Fare 

**cabin**:     Cabin 

**embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 
 
 
```{r warning=FALSE, message=FALSE}
library(caret)
library(dplyr)
library(ggplot2)
library(RCurl)
```

### Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### Removing NA.

```{r}
train<-train[, !(colnames(train) %in% c('name'))]
train <-train %>% na.omit()
test<-test[, !(colnames(test) %in% c('name'))]
test <- test %>% na.omit()
```

### Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### Creating GBM model

```{r}
gbmModel <- suppressMessages(train(formula, data=train, method="gbm"))
gbmModel
ggplot(varImp(gbmModel))
ggplot(gbmModel, metric = "Accuracy")
ggplot(gbmModel, metric = "Kappa")
```

### Predicting with the test data set

```{r}
Survived <- predict(gbmModel, test)
r<-cbind(Survived, test)
ggplot(r, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Test Data Predicted")

ggplot(train, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Training Data")

total<-rbind(train, r)
ggplot(total, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Total")
```

