---
title: "Titanic - Decision Trees with Parallel Processing"
output: 
  html_document: 
    keep_md: yes
---

### Variable Descriptions

**survival**:  Survival (0 = No; 1 = Yes) 

**pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**name**:      Name 

**sex**:       Sex 

**age**:       Age 

**sibsp**:     Number of Siblings/Spouses Aboard 

**parch**:     Number of Parents/Children Aboard 

**ticket**:    Ticket Number 

**fare**:      Passenger Fare 

**cabin**:     Cabin 

**embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 
 

```{r warning=FALSE, message=FALSE}
library(doParallel)
library(caret)
library(ParallelForest)
library(dplyr)
library(ggplot2)
library(rattle)
library(RCurl)
```

### 1. Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### 2. Removing names and NAs.

```{r}
train<-train[, !(colnames(train) %in% c('name'))]
train <-train %>% na.omit()
test<-test[, !(colnames(test) %in% c('name'))]
test <- test %>% na.omit()
```

### 3. Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 4. Random Forest training model with parallel processing

```{r}
set.seed(2020)
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
# The number of resampling iterations is 3.
fitControl <- trainControl(method = "cv", number = 3, allowParallel = TRUE)
```

```{r}
fit <- train(formula, data=train, method="rf", importance = TRUE, trControl=fitControl)
```

```{r}
stopCluster(cluster)
confusionMatrix.train(fit)
fit
# Accuracy
ggplot(fit, metric = "Accuracy")
# Kappa coefficient
ggplot(fit, metric = "Kappa")
```

### 5. Predicting with the test data set

```{r}
Survived<-predict(fit, test)
Survived
r<-cbind(Survived, test)
ggplot(r, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Test Data Predicted")

ggplot(train, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Training Data")

total<-rbind(train, r)
ggplot(total, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Total")
```
