---
title: "Titanic - Decision Trees"
output: 
  html_document: 
    keep_md: yes
---

### Variable Descriptions

**Survival**:  Survival (0 = No; 1 = Yes) 

**Pclass**:    Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd) 

**Name**:      Name 

**Sex**:       Sex 

**Age**:       Age 

**Sibsp**:     Number of Siblings/Spouses Aboard 

**Parch**:     Number of Parents/Children Aboard 

**Ticket**:    Ticket Number 

**Fare**:      Passenger Fare 

**Cabin**:     Cabin 

**Embarked**:  Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton) 

```{r warning=FALSE, message=FALSE}
library(rpart)
library(rattle)
library(dplyr)
library(RCurl)
```

### 1. Reading data

```{r}
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/train.csv')
train <- read.csv(text = url) 
url <- getURL('https://raw.githubusercontent.com/frankwwu/R-Knots/master/Titanic/test.csv')
test <- read.csv(text = url) 
```

### 2. Removing NAs

```{r}
train <- train %>% na.omit()
test <- test %>% na.omit()
```

### 3. Selecting features

```{r}
train$Survived <- factor(train$Survived)
formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked
```

### 4. Creating classification tree

```{r}
set.seed(200)
tree <- rpart(formula, data=train, method="class")
fancyRpartPlot(tree)
```

### 5. Cross-Validation 

To examine whether the tree model is over fitting, find the size of tree with the minimum error.

```{r}
plotcp(tree)
tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
printcp(tree)
```

### 6. Prune the tree
Prune back the tree to avoid overfitting the data. Typically, you will want to select a tree size that minimizes the cross-validated error, the xerror column printed by printcp( ).

```{r}
ptree<- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
fancyRpartPlot(ptree)
```

### 7. Prediction

```{r}
predict <- predict(ptree, test, type = "prob")
```

### 8. Visualizing the result

```{r}
test$Survived <- predict[,2]
ggplot(test, aes(Age, Fare, color=Survived)) + 
  geom_point(alpha = 0.5) +
  facet_grid(Pclass~Sex) +
  ggtitle("Predictiom with the Test Data ")
